---
title: "HW3_MasonHolcombe"
format: html
editor: visual
---

## Mason Holcombe

### Chapter 4

#### 3)

a\) Covariance = 2 and does not 0, not independent

b\) Covariance = 0, so X2 and X3 are independent

c\) Cov(x1,x3) = 0, cov(x2, x3) = 0, so (X1,X2) and X2 are independent

d\) 1/2 \* (cov(x1,x3) and cov(x2, x3)). Both of these are 0, so 1/2 \* 0 = 0. D is independent

e\) cov(x2,x2) = 5, cov(-5/2x1, x2) = -5/2 \* cov(x1,x2) = -5/2 \* -2 = 5, and cov(x2, x3) = 0. So the final covariance is 5 + 5 + 0 = 10 and this does not equal 0. So x2 and (e) are not independent.

#### 4)

a\)

```{r}
mu = c(2,-3,1)
sigma = matrix(c(1,1,1,
                 1,3,2,
                 1,2,2), byrow=T, ncol=3)
a = c(3, -2, 1)
```

```{r}
mean = t(a) %*% mu
var = t(a) %*% sigma %*% a
c(mean, var)
```

3x1 - 2x2 + x2 \~ N(13, 9)

#### 19)

a\) \~ $\chi_6^2$ chi2 with 6 df

b\)

-   p-variate normal with mu as mean and 1/20 \* sigma as covariance matrix

-   p-variate normal with 0 as mean and sigma as covariance matrix

c\) wishart distribution with 19 (20-1) df and sigma as the scale matrix

#### 29)

```{r}
air = read.table("T1-5.DAT")
x5 = air["V5"]
x6 = air["V6"]

X = cbind(x5, x6)
n = nrow(X)
```

a\)

```{r}
xbar = colMeans(X)
S = cov(X)
Sinv = solve(S)

result = numeric(n)
for (i in 1:n){
  xi = matrix(as.numeric(X[i, ]),
              ncol=1)
  result[i] = t(xi - xbar) %*% Sinv %*% (xi - xbar)
}
(result)
```

b\)

```{r}
cutoff = qchisq(0.5, df=2)
mean(result <= cutoff)
```

c\)

```{r}
results_sorted = sort(result)
theoretical_chi2 = qchisq((1:n - (1/2)) / n, df=2)

plot(theoretical_chi2, results_sorted)
abline(0,1,col='red')
```

#### 39)

```{r}
data39 = read.table("T4-6.DAT")
indep = data39["V1"]
supp = data39["V2"]
benev = data39["V3"]
conform = data39["V4"]
leader = data39["V5"]

gender = data39["V6"]
socio = data39["V7"]
```

a\)

Independence - NORMAL

```{r}
hist(indep$V1)
qqnorm(indep$V1)
```

Support -- NOT NORMAL

```{r}
hist(supp$V2)
qqnorm(supp$V2)
```

Benevolence -- NORMAL

```{r}
hist(benev$V3)
qqnorm(benev$V3)
```

Conform - NORMAL

```{r}
hist(conform$V4)
qqnorm(conform$V4)
```

Leader -- NOT NORMAL (SKEWED RIGHT)

```{r}
hist(leader$V5)
qqnorm(leader$V5)
```

b\)

```{r}
library("MVN")
```

```{r}
varnames = c("V1", "V2", "V3", "V4", "V5")
features = data39[ ,varnames]
```

```{r}
mvn_result = mvn(data = features, mvn_test = "mardia")
(mvn_result$multivariate_normality)

mvn_result2 = mvn(data = features, mvn_test = "hz")
(mvn_result2$multivariate_normality)
```
